### 2022-06-29 Report/manuscript.tex
First Final Version of the Report

    \subsection*{Implementation HOSVD}

@@

    \subsection*{Validation of HOSVD}

    The multilinear rank advantage is maximized when it comes to process sparce tensors since the dimmensions can be greatly reduced without losing of too much relevant information by analysing the profile of its multiples unfoldings.

    \begin{align}
        \mathcal{X} &\in \mathbb{C}^{8 \times 4 \times 10} \to \hat{\mathcal{X}} \in \mathbb{C}^{R_{1} \times R_{2} \times R_{3}}, \\ 
        \mathcal{X} &\in \mathbb{C}^{5 \times 5 \times 5} \to \hat{\mathcal{Y}} \in \mathbb{C}^{P_{1} \times P_{2} \times P_{3}},
    \end{align}

@@

    \subsection*{Implementation HOOI}

@@

    \subsection*{Validation of HOOI}

    The multilinear rank advantage is maximized when it comes to process sparce tensors since the dimmensions can be greatly reduced without losing of too much relevant information by analysing the profile of its multiples unfoldings.

    \begin{align}
        \mathcal{X} &\in \mathbb{C}^{8 \times 4 \times 10} \to \hat{\mathcal{X}} \in \mathbb{C}^{R_{1} \times R_{2} \times R_{3}}, \\ 
        \mathcal{X} &\in \mathbb{C}^{5 \times 5 \times 5} \to \hat{\mathcal{Y}} \in \mathbb{C}^{P_{1} \times P_{2} \times P_{3}},
    \end{align}

@@

@@

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.75\linewidth]{figs/hw11a1.png} \par 
        \caption{Convergence behavior of ALS algorithm.}
        \label{fig:hw11a1} 
    \end{figure}

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.75\linewidth]{figs/hw11a2.png} \par 
        \caption{Monter Carlo Experimento with 1000 runs for ALS algorithm.}
        \label{fig:hw11a2} 
    \end{figure}

@@

@@

@@

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.75\linewidth]{figs/hw12.png} \par 
        \caption{Monter Carlo Experiment with 1000 runs for TKPSVD algorithm.}
        \label{fig:hw12} 
    \end{figure}

@@

@@

@@

@@

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.75\linewidth]{figs/hw13.png} \par 
        \caption{Monter Carlo Experiment with 1000 runs for TTSVD algorithm.}
        \label{fig:hw13} 
    \end{figure}

>>>>

    \subsection*{Implementation and Validation of HOSVD}

    The HOSVD algorithm was implemented and validated according the file $"hosvd\_test.mat"$. The HOSVD algorithm is straighfoward  and its process is made by sucessive SVDs
    for each unfolding of the tensor $\mathcal{X}$. Each $\boldsymbol{U}_{n}$ matrix is formed by the most left autovectors of the SVD of the $n$th mode and the core tensor 
    $\mathcal{S}$ is made by the inverse multilinear transformation that describes the original tensor $\mathcal{X}$. Finally in the following table there is the NMSE between
    the validation files and the ones generated by my algorithm.

@@

    As we can see once more in the HOSVD there is also an scale ambiguity that makes impossible to recover the exact 
    $\boldsymbol{U}_{n}$ matrices without previous knowledge of the original structure. However, we can see that the reconstruction is perfect since the scale ambiguity cancel out each
    other when they are used to compose the estimation of the original tensor.

    \subsection*{HOSVD for Data Compression}

    It was asked to implement the truncated HOSVD and validate it with the file $"hosvd\_denoising.mat"$. However as far as I could see the file was not provided in the folder with the validation files.
    Nonetheless the truncated HOSVD was implemented by using the inspection of the SVD profiles of each unfolding of the tensor $\mathcal{X}$. The approximation was made by discarding the left singular vector of the SVD
    that were associated with eigenvalues which values were close to zero. In the context of my M.Sc. research this function was used to truncate the dimensions of the core tensor and its factors matrices and it was indeed possible
    to verify that some information can be lost when discarding eigenvalues but the order of loss can usually be controlled by the user.

@@

    \subsection*{Implementation and Validation of HOOI}

    The HOOI algorithm was implemented and validated according the file $"hosvd\_test.mat"$. The HOOI algorithm uses the HOSVD as initialization and procceds similarly to the HOSVD by taking sucessive
    SVDs of each unfolding of the original tensor $\mathcal{X}$. Each $\boldsymbol{U}_{n}$ matrix is formed by the most left autovectors of the SVD of the $n$th mode and the core tensor 
    $\mathcal{S}$ is made by the inverse multilinear transformation that describes the original tensor $\mathcal{X}$. Finally in the following table there is the NMSE between
    the validation files and the ones generated by my algorithm.

@@

    As we can see again in the HOOI there is also an scale ambiguity that makes impossible to recover the exact 
    $\boldsymbol{U}_{n}$ matrices without previous knowledge of the original structure. However, we can see that the reconstruction is perfect since the scale ambiguity cancel out each
    other when they are used to compose the estimation of the original tensor.

    \subsection*{HOOI for Data Compression}

    It was asked to implement the truncated HOOI and validate it with the file $"hosvd\_denoising.mat"$. However as far as I could see the file was not provided in the folder with the validation files.
    Nonetheless the truncated HOOI was implemented by using the previous full-rank HOSVD as initialization and by further inspecting the SVD profiles of each unfolding of the tensor $\mathcal{X}$. The approximation was made by discarding the left singular vector of the SVD
    that were associated with eigenvalues which values were close to zero.

@@

    Considering a third order tensor $\mathcal{X}$ obtained by the following expression

    \begin{align}
        \mathcal{X} = \sum^{R}_{r = 1} \boldsymbol{a}_{r} \circ \boldsymbol{b}_{r} \circ \boldsymbol{c}_{r} \in \mathbb{C}^{I_{1} \times I_{2} \times I_{3}},  
    \end{align}

    the ALS algorithm aims to solve the following estimation problem

@@

    and as we can see in the following table it is possible to verify the validation of the implemented ALS algorithm.

    \begin{table}[ht!]
        \centering
        \begin{tabular}{|l|l|l|l|}
        \hline
        NMSE($\mathcal{X}, \mathcal{\hat{X}}$) & NMSE($\boldsymbol{A}, \boldsymbol{\hat{A}}$) & NMSE($\boldsymbol{B}, \boldsymbol{\hat{B}}$) & NMSE($\boldsymbol{C}, \boldsymbol{\hat{C}}$) \\ \hline
        -312.1052 & +26.0794 & -0.2457 & +29.8936 \\ \hline
        \end{tabular}
    \end{table}

    As to be expected the reconstruction of the original tensor by the estimates of the factor matrices is perfect, but we can not say the same 
    for the individual estimations of the factor matrices. This is according to the theory since there is a intrinsic scale ambiguity to the ALS
    algorithm that makes impossible to perfectly estimate the factors without previous knowledge of its structure. However, when we use the estimated factors
    to reconstruct an estimation for the original tensor the scales ambiguities cancel out each other and then we obtain a perfect estimation of the tensor $\mathcal{X}$.

    \newpage
    \subsection*{Monte Carlo Experiment}

    In Figure \ref{fig:hw11a1} we took a snapshot of the ALS convergence behavior for a random experiment of the asked Monte Carlo Experiment.
    It is possible to see that for this specific experiment the algorithm took a bit more than 500 iterations to reach the desired convergence.
    For other experiments it could have taken even less than 50 iterations. Furthermore, if convergence speed is a problem the vanilla ALS could 
    also be modified to a newer version of the algorithm like the Rectified ALS or Enhanced Line Search (ELS) ALS.

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.60\linewidth]{figs/hw11a1.png} \par 
        \caption{Convergence behavior of ALS algorithm.}
        \label{fig:hw11a1} 
    \end{figure}

    In Figure \ref{fig:hw11a2} the Monte Carlo Experiment is used to draw the average behavior of the estimation of the factors 
    that generate a tensor $\mathcal{X}$ using the ALS algorithm. As we can observe the factor matrices does not show any improvements in its 
    estimation as the SNR gets bigger. This is just a consequence of the scale ambiguity problem that was commented in the previous section. Nonentheless, 
    we can also verify that at least for the curve that represents that tensor reconstruction the performance gets better as the variance of the noise that 
    corrupts the original tensor $\mathcal{X}$ decreases.

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.60\linewidth]{figs/hw11a2.png} \par 
        \caption{Monter Carlo Experimento with 1000 runs for ALS algorithm.}
        \label{fig:hw11a2} 
    \end{figure}

@@

    By using the tensor kronecker product operator it is possible to write the following tensor $\mathcal{X}$

@@

    and this tensor can be further decomposed by solving the following problem using the TKPSVD algorithm

    \begin{align}
        \left( \mathcal{A}^{(1)}_{j}, \mathcal{A}^{(2)}_{j}, \cdots, \mathcal{A}^{(d)}_{j} \right) = \underset{\mathcal{A}^{(1)}_{j}, \mathcal{A}^{(2)}_{j}, \cdots, \mathcal{A}^{(d)}_{j}}{\text{min}} \left| \left| \mathcal{X} - \sigma_{i} \mathcal{A}^{(d)}_{j} \otimes \mathcal{A}^{(d-1)}_{j} \otimes \cdots \otimes \mathcal{A}^{(1)}_{j} \right| \right|_{\text{F}},
    \end{align}

    and as we can see in the following table it is possible to verify the validation of the implemented TKPSVD algorithm.

@@

    As to be expected the reconstruction of the original tensor by the estimates of the factor matrices is perfect, but we can not say the same 
    for the individual estimations of the factor matrices. This is according to the theory since there is a intrinsic scale ambiguity to the TKPSVD
    algorithm that makes impossible to perfectly estimate the factors without previous knowledge of its structure. However, when we use the estimated factors
    to reconstruct an estimation for the original tensor the scales ambiguities cancel out each other and then we obtain a perfect estimation of the tensor $\mathcal{X}$.

    \newpage
    \subsection*{Monte Carlo Experiment}

    In Figure \ref{fig:hw12} the Monte Carlo Experiment is used to draw the average behavior of the TKPSVD. 
    As we can observe the factor matrices does not show any improvements in its estimation as the SNR gets bigger. This 
    is just a consequence of the scale ambiguity problem that was commented in the previous section. Nonentheless, we can also verify that at least for the curve that represents that tensor reconstruction the performance gets better as
    the variance of the noise that corrupts the original tensor $\mathcal{X}$ decreases.

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.85\linewidth]{figs/hw12.png} \par 
        \caption{Monter Carlo Experiment with 1000 runs for TKPSVD algorithm.}
        \label{fig:hw12} 
    \end{figure}

@@

    By using the tensor contraction operator it is possible to write the following tensor $\mathcal{X}$

@@

    and this tensor can be further decomposed by solving the following problem using the TTSVD algorithm

@@

    and it is also important to note that the TTSVD one depends of an additional input of the Tensor Train (TT) ranks. 
    In the validation experiment it was generated a tensor $\mathcal{X}$ that follows a TT structure of $R = (3,3,3)$ and the TTSVD algorithm 
    was used considering two different scenarios: The TT ranks $R = (3,3,3)$ and $R = (2,2,2)$. As we can see in the table the TTSVD algorithm performance
    is strongly associated with the correct input of the TT ranks, meaning that the previous knowledge of its values are a must to obtain satisfactory results 
    for this decomposition.

@@

    Finally, in Figure \ref{fig:hw13} we can confirm what was observed in the previous section. The correct knowledge of the TT ranks are essential to reach
    a decent performance for the TTSVD algorithm. Nonentheless, we can also verify that at least for the scenario $R = (3,3,3)$ the performance gets better as
    the variance of the noise that corrupts the original tensor $\mathcal{X}$ decreases.

    \begin{figure}[ht!]
        \centering 
        \includegraphics[width=0.65\linewidth]{figs/hw13.png} \par 
        \caption{Monter Carlo Experiment with 1000 runs for TTSVD algorithm.}
        \label{fig:hw13} 
    \end{figure}
